---
title: "Copy-number signatures and mutational processes in ovarian carcinoma"
output: 
    html_document:
        number_sections: false
        theme: simplex
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
suppressMessages(library(QDNAseq))
suppressMessages(library(flexmix))
suppressMessages(library(NMF))
suppressMessages(source("../main_functions.R"))
num_cores<-16
```

This document is compiled from an Rmarkdown file which contains all code necessary to reproduce the analysis for the accompanying manuscript ["Copy-number signatures and mutational processes in ovarian carcinoma"](https://www.biorxiv.org/content/early/2017/09/04/174201). Details on how to compile the document can be found in the  repository README: https://bitbucket.org/britroc/cnsignatures. Much of the code for signature analysis can be found in the main_functions.R and helper_functions.R files in the base directory of the repositoy. Each section below describes a different component of the analysis and all numbers and figures are generated directly from the underlying data on compilation.

##Data preprocessing

```{r child = 'CN_signature_identification_validation.Rmd'}
```

##Survival analysis

```{r child = 'Britroc_survival_estimation.Rmd'}
```

```{r child = 'CN_signature_survival_analysis.Rmd'}
```

##Association analysis

```{r child = 'CN_signature_association_analysis.Rmd'}
```

##Treatment response analysis
```{r child = 'CN_signature_relapse_analysis.Rmd'}
```
