---
title: "CN signature survival analysis"
author: "Geoff Macintyre"
output: html_document
---


Given that the BritROC study only enrolled patients with relapsed disease, it was necessary to left truncate the overall survival time. In addition, cases where the patient was not deceased were right censored. We combined the BritROC, PCAWG and TCGA cases and fit a Cox proportinal hazards model. We stratifed by study (BritROC, OV-US, OV-AU and TCGA).

###Compiling training and test cohorts
```{r echo=TRUE}
library(survival)
set.seed(seed)
samp_num<-0.7 #fraction of the data considered training

tcga_clin<-read.table("data/tcga_sample_info.tsv",stringsAsFactors = F,header=T,sep="\t")
pcawg_clin<-read.table("data/pcawg_sample_info.tsv",sep="\t",header=T,stringsAsFactors = F)
pcawg_cohort<-read.table("data/pcawg_cohort_info.tsv",header=T,sep="\t",stringsAsFactors = F)

#britroc survival data
pat_info<-samp_annotation[,c("Britroc_No","IM.JBLAB_ID","star_rating")]
pat_info<-samp_annotation[grepl("IM",samp_annotation$IM.JBLAB_ID),]
pat_info<-pat_info[order(pat_info$Britroc_No,pat_info$star_rating,decreasing = T),]
pat_info<-pat_info[!duplicated(pat_info$Britroc_No),]
pat_info<-pat_info[!pat_info$IM.JBLAB_ID%in%c("IM_91","IM_70"),]#remove misclassified samples
surv_dat<-read.table("data/britroc_survival_intervals.tsv",sep="\t",header=T,stringsAsFactors = F)
surv_dat<-surv_dat[!duplicated(surv_dat$TRIALNO),]
britroc_surv_dat<-merge(surv_dat,pat_info,by.x=1,by.y=1)
rownames(britroc_surv_dat)<-britroc_surv_dat$IM.JBLAB_ID
britroc_surv_dat<-britroc_surv_dat[colnames(sig_pat_mat_britroc_all)[colnames(sig_pat_mat_britroc_all)%in%rownames(britroc_surv_dat)],]
britroc_surv_dat<-britroc_surv_dat[!(is.na(britroc_surv_dat$INT_START)|is.na(britroc_surv_dat$INT_END)),]
britroc_surv_dat<-britroc_surv_dat[!(britroc_surv_dat$INT_START>britroc_surv_dat$INT_END),]
britroc_train<-rownames(britroc_surv_dat)
britroc_test<-rownames(britroc_surv_dat)[!rownames(britroc_surv_dat)%in%britroc_train]

#pcawg survival data
pdat<-pcawg_cohort[pcawg_cohort$tumor_wgs_aliquot_id%in%colnames(sig_pat_mat_pcawg),c("tumor_wgs_aliquot_id","icgc_donor_id","dcc_project_code")]
pdat<-merge(pdat,pcawg_clin[,c("icgc_donor_id","donor_vital_status","donor_survival_time","donor_interval_of_last_followup","donor_age_at_diagnosis")],by.x=2,by.y=1)
pdat<-pdat[!duplicated(pdat$tumor_wgs_aliquot_id),]
rownames(pdat)<-pdat$tumor_wgs_aliquot_id
pdat<-pdat[colnames(sig_pat_mat_pcawg),]
pcawg_surv_dat<-pdat
pcawg_os<-pcawg_surv_dat$donor_survival_time
pcawg_os[is.na(pcawg_os)]<-pcawg_surv_dat$donor_interval_of_last_followup[is.na(pcawg_os)]
pcawg_surv_dat$os<-pcawg_os
pcawg_event<-pcawg_surv_dat$donor_vital_status
pcawg_event<-plyr::revalue(pcawg_event,c("deceased"=1,"alive"=0))
pcawg_surv_dat$event<-pcawg_event
pcawg_train<-sample(rownames(pcawg_surv_dat),round(samp_num*nrow(pcawg_surv_dat)))
pcawg_test<-rownames(pcawg_surv_dat)[!rownames(pcawg_surv_dat)%in%pcawg_train]

#tcga survival data
tcga_surv_dat<-tcga_clin
tcga_surv_dat<-tcga_surv_dat[!is.na(tcga_surv_dat$age_at_initial_pathologic_diagnosis),]
tcga_surv_dat<-tcga_surv_dat[tcga_surv_dat$bcr_aliquot_barcode%in%colnames(sig_pat_mat_tcga),]
tcga_surv_dat<-tcga_surv_dat[!is.na(tcga_surv_dat$days_to_last_followup),]
tcga_survival_sigmat<-sig_pat_mat_tcga
tcga_os<-tcga_surv_dat$days_to_death
tcga_status<-!is.na(tcga_os)
tcga_surv_dat$status<-tcga_status
tcga_os[is.na(tcga_os)]<-tcga_surv_dat$days_to_last_followup[is.na(tcga_os)]
tcga_surv_dat$os<-tcga_os
tcga_train<-sample(rownames(tcga_surv_dat),round(samp_num*nrow(tcga_surv_dat)))
tcga_test<-rownames(tcga_surv_dat)[!rownames(tcga_surv_dat)%in%tcga_train]


combined_sig_pat_mat<-cbind(sig_pat_mat_britroc_all[,rownames(britroc_surv_dat)],
                            sig_pat_mat_pcawg[,rownames(pcawg_surv_dat)],
                            sig_pat_mat_tcga[,tcga_surv_dat$bcr_aliquot_barcode])

#threshold low exposure values to 2%
combined_sig_pat_mat[combined_sig_pat_mat<=0.02]<-0.02

combined_os_survival<-survival::Surv(time= c(britroc_surv_dat$INT_START,
                                   rep(1,length(pcawg_os)),
                                   rep(1,length(tcga_os))),
                           time2=c(britroc_surv_dat$INT_END,
                                   pcawg_surv_dat$os,
                                   tcga_surv_dat$os),
                           event=c(!britroc_surv_dat$STATUS,
                                   as.numeric(pcawg_surv_dat$event)==1,
                                   tcga_surv_dat$status))
survival_ids<-c(rownames(britroc_surv_dat),rownames(pcawg_surv_dat),rownames(tcga_surv_dat))

sig_data<-data.frame(t(combined_sig_pat_mat),
                     study=c(rep("britroc",nrow(britroc_surv_dat)),
                             pcawg_surv_dat$dcc_project_code,rep("TCGA",nrow(tcga_surv_dat))),
                     age=c(britroc_surv_dat$AGE,
                           pcawg_surv_dat$donor_age_at_diagnosis,
                           tcga_surv_dat$age_at_initial_pathologic_diagnosis),
                     stringsAsFactors = F)
#stratify by age
sig_data$age.cat<-car::recode(as.numeric(sig_data$age), "lo:39=1; 40:44=2; 45:49=3; 50:54=4; 55:59=5; 60:64=6; 65:69=7; 70:74=8; 75:79=9; 80:hi=10")
```

As the signatures sum to 1, in order to perform survival analysis the exposures had to be normalised using the signature with the lowest variance:

```{r}
#determine signature with lowest variance
which.min(apply(combined_sig_pat_mat,1,var))

#normalise signature exposures
sig_data_unorm<-sig_data
sig_data[,1:nsig]<-apply(sig_data[,1:nsig],2,function(x){log2(mapply("/",x,sig_data[,5]))})

#split data into training and test cohorts
sig_data_train<-sig_data[survival_ids%in%c(britroc_train,pcawg_train,tcga_train),]
sig_data_test<-sig_data[survival_ids%in%c(britroc_test,pcawg_test,tcga_test),]
combined_os_survival_train<-combined_os_survival[survival_ids%in%c(britroc_train,pcawg_train,tcga_train)]
combined_os_survival_test<-combined_os_survival[survival_ids%in%c(britroc_test,pcawg_test,tcga_test)]
```

###Fitting a cox-proportional hazard model (training)
Here we fit a cox proportional hazards model to predict survival with the normlised signature exposures as covariates, stratified by study and age.
```{r coxph, echo=TRUE}
os_coxph<-coxph(combined_os_survival_train ~ s1+s2+s3+s4+s6+s7 + strata(study, age.cat,na.group=T), data = sig_data_train)
summary(os_coxph)

#test for proportional hazards
cox.zph(os_coxph)
```

Note: exp(coef) is the Hazard ratio.

###Testing the cox-proportional hazard model (test)
Here we tested the ability of the cox proportional hazards model to predict overall survival using the concordance index. 
```{r echo=TRUE}
#predict survival using coxph model
train_predict<-predict(os_coxph,type="risk")
test_predict<-predict(os_coxph,newdata=sig_data_test,type="risk")

perf <- survcomp::concordance.index(x = test_predict, surv.time = combined_os_survival_test[,"stop"],
                          surv.event = combined_os_survival_test[, "status"], method = "noether",na.rm = TRUE)
perf[1:5]
```

###Fitting cox-proportional hazard model (all data)
We then fit a second cox proportional hazards model to the combined test and training data.
```{r echo=TRUE}
os_coxph<-coxph(combined_os_survival~ s1+s2+s3+s4+s6+s7 + strata(study, age.cat,na.group=T), data = sig_data)
summary(os_coxph)

#check for non-proportional hazards
cox.zph(os_coxph)

#Compute hazard ratio table
predicted_survival<-predict(os_coxph,type="risk")

HR <- round(exp(coef(os_coxph)), 2)
CI <- round(exp(confint(os_coxph)), 2)
P <- round(coef(summary(os_coxph))[,5], 3)
colnames(CI) <- c("Lower", "Higher")
hazard_table <- as.data.frame(cbind(HR, CI, P))
hazard_table<-cbind(Signature=row.names(hazard_table),hazard_table)
knitr::kable(hazard_table)
```

###Unsupervised clustering of patients using signature exposures
To group patients based on their copy-number signature exposures we used the [NBClust](https://cran.r-project.org/package=NbClust) package to determine the number of clusters present in the data and cluster the patients.

```{r}
mydata<-t(combined_sig_pat_mat[,])
#determine number of clusters
fit <- NbClust::NbClust(mydata,method="ward.D2")

#visualise signature exposures within clusters
sig_data$clust=fit$Best.partition

pdat<-data.frame(t(combined_sig_pat_mat))
pdat$clust<-fit$Best.partition
pdat$sample<-rownames(pdat)
pdat<-reshape2::melt(pdat,id.vars=c(8,9))
pdat$variable<-plyr::revalue(pdat$variable,
                            c(s1=1,s2=2,s3=3,s4=4,s5=5,s6=6,s7=7))
ggplot(pdat,aes(x=variable,y=value))+geom_boxplot()+geom_jitter(alpha=0.1)+facet_grid(clust ~ .)+my_theme+xlab("Signature")+ylab("Exposure")
```

###Clustering membership plot
```{r, fig.height=2}
pdat$sample<-factor(pdat$sample,levels=names(predicted_survival)[order(predicted_survival,decreasing=T)])
pdat$clust<-factor(pdat$clust,levels=3:1)
clust_plot<-ggplot(pdat,aes(y=clust,x=sample))+geom_tile()+my_theme+theme(axis.text.x =element_blank(),axis.ticks.x = element_blank())+
  xlab("")+ylab("")+scale_fill_discrete()+theme(legend.position = "none")
clust_plot
```

###Fitting cox-proportional hazard model for clusters
We also fit a cox-proportional hazards model using the clusters as covariates.
```{r echo=TRUE}
sig_data$clust<-factor(sig_data$clust,levels=3:1)
os_coxph<-coxph(combined_os_survival~ clust + strata(study, age.cat), data = sig_data)
summary(os_coxph)
cox.zph(os_coxph)
```

While this model showed significant predictive preformance, the exposures across the clusters (above) reveal that cluster 2 is based primarily on signal from signature 1 exposures, and thus a cluster based model does not add information above and beyond the signature model.

###Visualising cox-proportional hazard model
```{r echo=FALSE}
#plot all survival
pdat_survsig<-reshape2::melt(cbind(sig_pat_mat_britroc,sig_pat_mat_pcawg,sig_pat_mat_tcga))
colnames(pdat_survsig)<-c("Signature","Patient","Exposure")
pdat_survsig<-pdat_survsig[pdat_survsig$Patient%in%names(predicted_survival),]
pdat_survsig$Patient<-factor(pdat_survsig$Patient,levels=names(predicted_survival)[order(predicted_survival,decreasing=T)])
pdat_survsig$Patient<-factor(pdat_survsig$Patient)

p1<-ggplot(pdat_survsig,aes(x=Patient,y=as.numeric(Exposure),fill=Signature))+geom_bar(stat="identity")+
      scale_fill_manual( values = cbPalette)+my_theme+
      theme(axis.text.x = element_blank(),axis.ticks.x=element_blank(),
            axis.text=element_text(size=5),axis.title=element_text(size=7),
            legend.text = element_text(size = 7),legend.title=element_text(size=7),
            legend.position='none',
            plot.margin = margin(0, 0, 0, 0, "cm"),panel.background = element_blank())+xlab("")+
      ylab("")

p2<-ggplot(pdat_survsig,aes(x=Patient,y=Exposure,colour=Signature,group=Signature))+
      geom_smooth(se=F,lwd=0.3,method="lm")+
      theme_bw()+
      scale_colour_manual(values = cbPalette)+
      theme(axis.text.x = element_blank(),axis.ticks.x=element_blank(),
            axis.text=element_text(size=5),axis.title=element_text(size=7),
            legend.text = element_text(size = 7),legend.title=element_text(size=7),
            legend.position='none',
            plot.margin = margin(-0.1, 0, 0, 0, "cm"))+xlab("")+
      ylab("")
cowplot::plot_grid(p1,clust_plot,p2,align="v",ncol=1,rel_heights = c(0.45,0.1,0.45))
```